\documentclass{report}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage[titles]{tocloft}
\usepackage{tikz}
\usepackage{xcolor}

\usepackage{multicol}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{pdfpages}
\usepackage{bm}
%\usepackage{bbm}
%\usepackage{biblatex}
%\addbibresource{bib.bib}

%hyperref should be last apparently
\usepackage{hyperref}

\renewcommand\cftsecdotsep{\cftdot}
\renewcommand\cftsubsecdotsep{\cftdot}
\renewcommand\epsilon{\varepsilon}

% Starts a new paragraph without indentation
% and with an empty line between paragraphs
\newcommand*{\newpar}{\par\vspace{\baselineskip}\noindent}
\newcommand{\trans}{\twoheadrightarrow}
\newcommand{\ttt}[1]{\texttt{#1}}
\newcommand{\tbf}[1]{\textbf{#1}}

\pagestyle{fancy} %allows headers

\lhead{Emma Bach}
\rhead{\today}


\begin{document}
\newtheorem{lemma}{Lemma}[section]
\newtheorem{theorem}[lemma]{Theorem}

\theoremstyle{definition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{example}[lemma]{Example}
%
%
%
\include{title}
\tableofcontents
\thispagestyle{fancy}
%
%
%
\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
These notes are not based on any particular lecture I attended. Instead, they were created during self-study of theoretical topics related to type theory, functional programming and the $\lambda$-calculus.
\newpar
My primary sources are ``Types and Programming Languages'' by Benjamin C. Pierce \cite{types} and the ``Handbook of Logic in Computer Science'', edited by Samson Abramsky, Dov  Gabbay and Thomas Maibaum, in particular volume 2, which deals with computational structures \cite{handbookvol2}.  Additional sources I found helpful include the video ``What is plus times plus?'' by YouTube channel 2Swap, which is a wonderful introduction to the $\lambda$-calculus \cite{2swap}.
\newpar
The first few chapters correspond to \cite{types}, which tends to use informal, intuitive definitions, while the latter are generally from \cite{handbookvol2}, which is very formal and rigourous. These differences will be noticable in the notes too.
%
\chapter{Untyped Systems}
\section{Untyped Arithmetic Expressions}
This chapter serves as an introduction of many concepts related to formal languages, such as abstract syntax, inductive definitions, evaluations, and the modeling of run-time errors.
\newpar
In this first section, we will examine a simple language containing the following elements:
\begin{itemize}
 \item the boolean constants \texttt{true} and \texttt{false}
 \item conditional expressions \texttt{if $t_1$ then $t_2$ else $t_3$}
 \item the numeric constant $0$
 \item the arithmetic operators \texttt{succ} and \texttt{pred}
 \item a 1-ary function \texttt{iszero}
\end{itemize}
We summarize these contents by writing them as the following grammar:
\begin{align*}
 t ::= &\\
 &\ttt{true}\\
 &\ttt{false}\\
 &\ttt{if t then t else t}\\
 &\ttt{0}\\
 &\ttt{succ t}\\
 &\ttt{pred t}\\
 &\ttt{iszero t}
\end{align*}
The first line $t ::=$ declares that we are defining a set of \tbf{terms}, and that we are using the letter $t$ to refer to these terms. Each line that follows gives a possible \tbf{syntactic form} for terms. The letter $t$ does not refer to any specific term - whenever $t$ appears, we may substitute any term. A symbol like $t$ is called a \tbf{metavariable}, the ``meta'' referring to the fact that it is not a variable within the language we are defining (also known as the \tbf{object language}), but rather in the notation that we are defining the language in (the \tbf{metalanguage}).
\newpar
This ``meta-'' naming scheme comes from \tbf{metamathematics}, which is the field of logic that deals with the mathematical properties of formal systems of mathematical reasoning, such as programming languages. The collection of true statements we can make about such a system is known as the \tbf{metatheory} of the system.
\newpar
A \tbf{program} in the given language is just a term belonging to it. If we additionally assign semantical meaning to the terms we defined, we can evaluate terms belonging to our system, by \tbf{reducing} them to another term belonging to the language. For example, the term \ttt{if false then 0 else 1;} evaluates to \ttt{1}, and the term \ttt{iszero(pred(succ(0)))} evaluates to \ttt{true}. Note that parentheses are added for readability, but often not included in the formal description of the language. Also note that, in our language, all natural numbers are formally represented as repeated applications of \ttt{succ} to \ttt{0}. However, for brevity, arabic numerals will be used to represent these terms. For example, whenever we use the term \ttt{2}, the equivalent term actually used in the language would be \ttt{succ(succ(0))}.
\newpar
A more formal mathematical definition of our language would use induction to define the set $\mathcal{T}$ of all terms:
\begin{definition}
The set of terms is the smallest set $\mathcal{T}$ such that:
\begin{enumerate}
 \item $\{\ttt{true}, \ttt{false}, \ttt{0}\} \subseteq \mathcal{T}$
 \item if $t_1 \in \mathcal{T}$, then $\{\ttt{succ $t_1$, pred $t_1$, iszero $t_1$}\}\subseteq \mathcal{T}$
 \item if $t_1, t_2, t_3 \in \mathcal{T}$, then $\ttt{if $t_1$ then $t_2$ else $t_3$} \in \mathcal{T}$
\end{enumerate}

\end{definition}
%
\chapter{Formal Term Rewriting Systems}
\section{Abstract Reduction Systems}
\begin{definition}
\hphantom{}
\begin{enumerate}
 \item An \textbf{abstract reduction system} is a structure $\mathcal{A} = \langle A, \{\rightarrow_{\alpha}\}_{\alpha \in I}\rangle$, consisting of a set of terms $A$ and a set of binary relations $\rightarrow_{\alpha} \in A^2$, called the \textbf{reduction relations} or the \textbf{rewrite relations}. As shorthand, the relation $\rightarrow_{\alpha}$ is occasionally written simply as $\alpha$. An abstract reduction system with only one relation $\rightarrow$ is also known as a \textbf{replacement system} or a \textbf{transformation system}.
 \item If $a \rightarrow_\alpha b$, then $b$ is known as a \textbf{one step ($\alpha$-) reduct} of $a$, meaning that $b$ can be created by applying the reduction $\rightarrow_\alpha$ to $a$ once.
 \end{enumerate}
\end{definition}
%
Note that we let $\rightarrow_\alpha$ be any arbitrary relation, without additional demands of transitivity, reflexivity, or anything else of the sort.
\begin{example}
As an example, we could consider a term rewriting system that reduces sums of natural numbers. To write this system as an abstract reduction system, we can define $A$ and $\rightarrow$ as follows:
\begin{itemize}
 \item Let $A$ be the smallest set of terms such that:
 \begin{itemize}
 \item Every natural number $n \in \mathbb{N}$ is a term.
 \item If $t_1$ and $t_2$ are terms, then $t_1 + t_2$ is a term.
 \end{itemize}
 \item We define the following reduction relations $\to_0$ and $\to_1$:
 \begin{itemize}
  \item If $n$ and $m$ are natural numbers and $k = n+m$, then $n + m \rightarrow_0 k$
  \item If $t_1 \to_0 k$, then $t_1 + t_2 \to_{1} k + t_2$
 \end{itemize}
\end{itemize}
We could now apply these reduction rules to reduce a sum:
\begin{equation*}
 10+5+2+7 \to_1 15 + 2 + 7 \to_1 17 + 7 \to_0 24
\end{equation*}
The term $15+2+7$ is a one step $1$-reduct of the term $10+5+2+7$, while the term $17+7$ is a one-step $1$-reduct of $15+2+7$ and therefore a two step $1$-reduct of $10+5+2+7$.
\end{example}
%
\begin{definition}
Given a reduction relation $\to_{\alpha}$, we are often interested in investigating new relations created by extending $\to_{\alpha}$. Some relations that are of interest to us include:
\begin{itemize}
 \item The \textbf{transitive reflexive closure} $\trans_\alpha$, defined as the smallest transitive relexive relation containing $\to_\alpha$. To give a more explicit definition, we have $a \twoheadrightarrow_\alpha b$ iff:
  \begin{enumerate}
   \item $a \to_\alpha b$, or
   \item $a = b$, or
   \item $\exists n : \exists c_1, \hdots, c_n : a \to_\alpha c_1 \to_\alpha \hdots \to_\alpha c_n \to_\alpha b$
  \end{enumerate}
  In most of the literature, this closure would be denoted as $\to_\alpha^*$. However, the book uses $\trans_\alpha$ because it makes transitive diagrams involving the relation more legible.\
  \item The \textbf{convertibility relation} $=_\alpha$ generated by $\to_\alpha$, defined as the smallest equivalence relation containing $\to_\alpha$. Once again, to give a more explicit definition:
  \begin{center}
   We have $a =_\alpha b$ iff. $\exists u_1, \hdots, u_n \in A$ such that $a = u_1$, $b = u_n$,\\ and for each $u_i$, we have either $u_i \to_\alpha u_{i+1}$ or $u_{i+1} \to_\alpha u_i$.
  \end{center}
  Note that even though $\trans_\alpha$ is the smallest transitive reflexive relation that contains $\to_\alpha$ and $=_\alpha$ is the smallest equivalence (transitive, reflexive, and symmetric) relation that contains $\to_\alpha$, $=_\alpha$ is \textbf{not} equivalent to the symmetric closure of $\trans_\alpha$, as can be seen in Example \ref{ex:symmetricclosure}.
\end{itemize}
%
\begin{example}
\label{ex:symmetricclosure}
Let $\to_\alpha = \{(a,c), (b,c), (c,d)\}$. Then:
\begin{enumerate}
 \item $\trans_\alpha = \{(a,a), (b,b), (c,c), (d,d), (a,c), (a,d), (b,c), (b,d), (c,d)\}$
 \item The symmetric closure of $\trans_\alpha$ would be\\ $\trans_\alpha^s = \{(a,a), (b,b), (c,c), (d,d), (a,c), (c,a), (a,d), (d,a), (b,c), (c,b), (b,d), (d,b), (c,d), (d,c)\}$. Note that this relation is no longer transitive, since it includes $(a,c)$ and $(c,b)$, but not $(a,b)$.
 \item $=_\alpha$ would include every single pair of these elements.
\end{enumerate}
In short, if $a \to_\alpha c \leftarrow_\alpha b$, then we have $a =_\alpha b$, but not necessarily $a \trans_\alpha^s b$.
\end{example}
%
\end{definition}
%
%
%
%
\bibliography{refs}{}
\bibliographystyle{plain}
\end{document}
